---
title: "lala"
author: "Sandra Fogg"
date: "2/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}

library(tibble)
library(fable)
library(janitor)
library(forecast)
library(janitor)
library(tidyverse)
library(tsibble)
library(paletteer)
library(lubridate)
library(tidyverse)
library(here)
library(feasts)
library(sf)
library(tmap)
library(mapview)
```

```{r}
us_renew <- read_csv("data/renewables_cons_prod.csv") %>% 
  clean_names()
```

Make description all lowercase
Only keep observations where description variable contains the string pattern "consumption"
Remove any observations where description variable contains "total"

```{r}
renew_clean <- us_renew %>% 
  mutate(description = str_to_lower(description)) %>% 
  filter(str_detect(description, pattern = "consumption")) %>% 
  filter(!str_detect(description, patter = "total"))
```

convert column to date
make it understood as time series associated data frame
```{r}
renew_date <- renew_clean %>% 
  mutate(yr_mo_day = lubridate::parse_date_time(yyyymm, "ym")) %>% 
  mutate(month_sep = yearmonth(yr_mo_day)) %>% 
  mutate(value = as.numeric(value)) %>% 
  drop_na(month_sep, value)
```

make version with month and year in separate columns
```{r}
renew_parsed <- renew_date %>% 
  mutate(month = month(yr_mo_day, label = TRUE)) %>%
  mutate(year = year(yr_mo_day))

renew_parsed
  
```

look at it
```{r}
renew_gg <- ggplot(data = renew_date, aes(x = month_sep, y = value)) + 
  geom_line(aes(color = description))

renew_gg
```

updating colors with paletteer

```{r}
renew_gg + scale_color_paletteer_d("ggthemr::lilac")
```

coerce renew_parsed to tsibble
```{r}
renew_ts <- as_tsibble(renew_parsed, key = description, index = month_sep)
```

look at time series a few ways
```{r}
renew_ts %>%
  autoplot(value)


#renew_ts %>% gg_subseries(value) # didn't work

#renew_ts %>% gg_season(value)



#make in gg plot!



ggplot(data = renew_parsed, aes(x = month, y = value, group = year)) +
  geom_line(aes(color = year )) +
  facet_wrap(~description,
             ncol = 1,
             scales = "free",
             strip.position = "right")
```

look at hydroelectric energy consumptions
```{r}
hydro_ts <- renew_ts %>% 
  filter(description == "hydroelectric power consumption")


hydro_ts %>% autoplot(value)
# sub series
# season

ggplot(hydro_ts, aes(x = month, y = value, group = year))+
  geom_line(aes(color = year))
```

what if i want the quartely average consuption for hydro

```{r}
hydro_quarterly <- hydro_ts %>% 
 index_by(year_qu = ~(yearquarter(.))) %>% 
  summarize(avg_consumption = mean(value)) # not working?
```

# Decompose that hydro_ts need the feast package!!!
```{r}
#dcmp <- hydro_ts %>% 
#  model(STL(value ~ season(window = 5)))
#components(dcmp) %>% 
#  autoplot()

#hist(components(dcmp)$remainder)

hydro_ts %>% 
  ACF(value) %>% 
  autoplot()

```
```{r}
### DANGER DANGER

hydro_model <- hydro_ts %>% 
  model(
    ARIMA(value)
  ) %>% 
  fabletools::forecast(h = "4 years")

hydro_model %>% autoplot(filter(hydro_ts, year(month_sep) > 2010))


# didnt work
```

make a world map
```{r}
world <- read_sf(dsn = here("data", "TM_WORLD_BORDERS_SIMPL-0.3-1"),
                 layer = "TM_WORLD_BORDERS_SIMPL-0.3")
```

